Modeling Ionic Polymer-Metal Composites\\ with Space-Time Adaptive Multimesh \emph{hp}-FEM

### 2010-10-19 model.tex
Pavel's suggestions, improving the results section

% IF YOU WANT TO CITE THIS WORK< YOU HAVE TO SAY WHY WE ARE DOING 
% THE DERIVATION AGAIN -- OTHERWISE IT IS CONTRAPRODUCTIVE. SO 
% I COMMENTED THIS OUT.
%An implicit version of the weak form of PNP was derived in~\cite{pugal2010spie}.
To make our results easily reproducible, we present a rather detailed
derivation of weak forms of Eqs.~\eqref{eq:nernst-planck} and 
\eqref{eq:poisson}, as well as formulas for the Jacobian matrix and residual 
vector that are used in actual computations.
To facilitate the derivation, we introduce two additional constants 
\begin{eqnarray}
  K & = &z \mu F,\\
  L&=&\frac{F}{\varepsilon}.
  \label{eq:KL}
\end{eqnarray}
After using Eq.~\eqref{eq:rho} and the constants $K$ and $L$, 
Eq.~\eqref{eq:nernst-planck} and Eq.~\eqref{eq:poisson} become
\begin{eqnarray}
  \frac{\partial C}{\partial t}-D\nabla^2 C-K\nabla\cdot \left(C\nabla\phi\right)&=&0,\label{eq:nernst-planck-2}\\
  -\nabla^2\phi-L\left(C-C_{0}\right)&=&0.\label{eq:poisson-2}
\end{eqnarray}
The boundary condition Eq.~\eqref{eq:nernst-planck-boundary} now has the form
\begin{equation}
  -D\frac{\partial C}{\partial n}-KC\frac{\partial\phi}{\partial n}=0.
  \label{eq:nernst-planck-boundary-2}
\end{equation}
As the second derivatives of both $C$ and $\phi$ are present in the 
equations, the appropriate function space for them is the Sobolev space 
$V=H^1\left(\Omega\right)$, where 
$H^1\left(\Omega\right)=\left\{v\in L^2\left(\Omega\right);\ \nabla v \in \left[L^2\left(\Omega\right)\right]^2\right\}$.

@@

Hermes\footnote{http://hpfem.org/hermes} is a free and open-source C++ library 
that implements higher-order finite elements approximations and adaptive $hp$-FEM.
It supports 8 different adaptivity modes -- three isotropic and five anisotropic. 
The isotropic refinements are
\emph{h}-isotropic (H\_ISO), \emph{p}-isotropic (P\_ISO), \emph{hp}-isotropic (HP\_ISO).
Anisotropic refinement modes are
\emph{h}-anisotropic (H\_ANISO),
\emph{hp}-anisotropic-\emph{h} (HP\_ANISO\_H), \emph{p}-anisotropic (P\_ANISO),
\emph{hp}-anisotropic-p (HP\_ANISO\_P), and \emph{hp}-anisotropic (HP\_ANISO).
The eight adaptivity modes are summarized in Fig. \ref{fig:candlist}.

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=0.7\columnwidth]{cand_list_quads}
  \caption{\label{fig:candlist} Refinement candidates for every
  refinement mode for quad type elements.}
  \end{centering}
\end{figure}
Note that triangular elements do not support anisotropic refinements.
Due to the large number of refinement options, classical error estimators 
that provide a constant error estimate per element, cannot be used to 
guide automatic \emph{hp}-adaptivity. 
For this, one needs to know the shape of the approximation error.
Hermes uses a pair of approximations with different orders of accuracy 
to obtain this information: coarse mesh solution and fine mesh solution. 
The initial coarse mesh is read from the mesh file, and the initial 
fine mesh is created through its global refinement both in \emph{h}
and \emph{p}. The fine mesh solution is the approximation of 
interest both during the adaptive process and at the end of computation. 
Global orthogonal projection of the fine mesh solution on the coarse mesh 
is used to extract the low-order part from the reference solution.
The adaptivity algorithm is guided by the difference between the 
reference solution and its low-order part [ADD CITATION SOLIN-SIMKO].
Note that this approach to automatic adaptivity is PDE-independent
and thus naturally applicable to a large variety of multiphysics 
coupled problems.

>>>>

% IF YOU WANT TO CITE THIS WORK< YOU HAVE TO SAY WHY WE ARE DOING 
% THE DERIVATION AGAIN -- OTHERWISE IT IS CONTRAPRODUCTIVE. SO 
% I COMMENTED THIS OUT.
% I guess it can be left out as the work in SPIE is quite preliminary
% and not that complete
%An implicit version of the weak form of PNP was derived in~\cite{pugal2010spie}.
To make our results easily reproducible, we present a rather detailed
derivation of weak forms of Eqs.~\eqref{eq:nernst-planck} and 
\eqref{eq:poisson}, as well as formulas for the Jacobian matrix and residual 
vector that are used in actual computations.
To facilitate the derivation, we introduce two additional constants 
\begin{eqnarray}
  K & = &z \mu F,\\
  L&=&\frac{F}{\varepsilon}.
  \label{eq:KL}
\end{eqnarray}
After using Eq.~\eqref{eq:rho} and the constants $K$ and $L$, 
Eq.~\eqref{eq:nernst-planck} and Eq.~\eqref{eq:poisson} become
\begin{eqnarray}
  \frac{\partial C}{\partial t}-D\nabla^2 C-K\nabla\cdot \left(C\nabla\phi\right)&=&0,\label{eq:nernst-planck-2}\\
  -\nabla^2\phi-L\left(C-C_{0}\right)&=&0.\label{eq:poisson-2}
\end{eqnarray}
The boundary condition Eq.~\eqref{eq:nernst-planck-boundary} now has the form
\begin{equation}
  -D\frac{\partial C}{\partial n}-KC\frac{\partial\phi}{\partial n}=0.
  \label{eq:nernst-planck-boundary-2}
\end{equation}
As the second derivatives of both $C$ and $\phi$ are present in the 
equations, the appropriate function space for them is the Sobolev space 
$V=H^1\left(\Omega\right)$, where 
$H^1\left(\Omega\right)=\left\{v\in L^2\left(\Omega\right);\ \nabla v \in \left[L^2\left(\Omega\right)\right]^2\right\}$.

@@

Hermes\footnote{http://hpfem.org/hermes} is a free and open-source C++ library 
that implements higher-order finite elements approximations and adaptive $hp$-FEM.
It supports 8 different adaptivity modes -- three isotropic and five anisotropic. 
The isotropic refinements are
\emph{h}-isotropic (H\_ISO), \emph{p}-isotropic (P\_ISO), \emph{hp}-isotropic (HP\_ISO).
Anisotropic refinement modes are
\emph{h}-anisotropic (H\_ANISO),
\emph{hp}-anisotropic-\emph{h} (HP\_ANISO\_H), \emph{p}-anisotropic (P\_ANISO),
\emph{hp}-anisotropic-p (HP\_ANISO\_P), and \emph{hp}-anisotropic (HP\_ANISO).
The eight adaptivity modes are summarized in Fig.~\ref{fig:candlist}.

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=0.7\columnwidth]{cand_list_quads}
  \caption{\label{fig:candlist} Refinement candidates for every
  refinement mode for quad type elements.}
  \end{centering}
\end{figure}
Note that triangular elements do not support anisotropic refinements.
Due to the large number of refinement options, classical error estimators 
that provide a constant error estimate per element, cannot be used to 
guide automatic \emph{hp}-adaptivity. 
For this, one needs to know the shape of the approximation error.
Hermes uses a pair of approximations with different orders of accuracy 
to obtain this information: coarse mesh solution and fine mesh solution. 
The initial coarse mesh is read from the mesh file, and the initial 
fine mesh is created through its global refinement both in \emph{h}
and \emph{p}. The fine mesh solution is the approximation of 
interest both during the adaptive process and at the end of computation. 
Global orthogonal projection of the fine mesh solution on the coarse mesh 
is used to extract the low-order part from the reference solution.
The adaptivity algorithm is guided by the difference between the 
reference solution and its low-order part~\cite{solin2010pde}.
Note that this approach to automatic adaptivity is PDE-independent
and thus naturally applicable to a large variety of multiphysics 
coupled problems.

### 2010-10-19 results.tex
Work on the first two pages.
Editing sections 1 - 4.
Pavel's suggestions, improving the results section

\section{Results}

\begin{figure}
  \begin{centering}
  \includegraphics[width=.8\columnwidth]{mesh}
  \caption{\label{fig:mesh} Initial coarse mesh (a),
  	half refined mesh (b) and refined mesh (c). The coarse mesh
	and refined mesh were used in the initial calculations, the latter one
	in case of \emph{p}-adaptivity (including HP\_ANISO\_P). The half-refined mesh was
	used later to optimize the \emph{hp}-adaptive refinement solutions.}
  \end{centering}
\end{figure}
Calculation for all the refinement modes were performed
in both single-mesh and multi-mesh configurations. 
The following numerical results were recorded for each 
refinement mode: converged relative error, cumulative CPU
time, and the problem size in terms of number
of degrees of freedoms (NDOFs) at each time step. 
Two types of initial meshes were used --- in case of only \emph{p}-adaptivity,
more refined mesh was used (Fig.~\ref{fig:mesh}~(c)) to ensure
the error convergence.
When the element size refinement
was also enabled (all \emph{h}/\emph{hp} refinement modes), very coarse initial mesh
was used (see Fig.~\ref{fig:mesh}~(a)) to let the adaptivity
algorithm find the most optimal mesh. So coarse initial mesh might not be
suitable for all the practical applications, as will be demonstrated in the
end of this section, however, it provides a good insight into the
adaptivity performance of Hermes.
In both cases, the initial mesh was loaded at each time step of the
calculation.

\begin{figure}
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{cphi}
  \caption{\label{fig:cphi} Concentration $C$
  and voltage $\phi$ at two different time steps
  (HP\_ANISO adaptivity was used). This is a 2D solution shown
  as in 3D, where the height indicates the values of $C$ and $\phi$.}
  \end{centering}
\end{figure}
Example of the solution at $t=0.7\ s$ and $t=3.0\ s$ 
with refinement mode HP\_ANISO is shown
in Fig.~\ref{fig:cphi}. The time $t=0.7\ s$ was chosen because
by that time, some ionic migration has already taken place, i.e.
concentration gradient near the $\partial\Omega_1$ and
$\partial\Omega_3$ has formed. The automatic mesh refinements
at different time steps are clearly visible in the figure --- especially
near the top boundary, where the concentration gradient is
moving in time (as was also seen in Fig.~\ref{fig:comsol-conc-volt}).

@@

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{singlemulti_dof}
  \caption{\label{fig:singlemultidof} NDOFs in case 
  of single-mesh and multi-mesh solutions with HP\_ISO
  and HP\_ANISO refinements. (Notice log Y scale)}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{singlemulti_cpu}
  \caption{\label{fig:singlemulticpu} CPU times in case
  of single-mesh and multi-mesh solutions with HP\_ISO
  and HP\_ANISO refinements. (Notice log Y scale)}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{poly}
  \caption{\label{fig:poly} Polynomial degree space
  for $C$ and $\phi$ at $t=0.7\ s$ The color indicates
  the maximum polynomial degree of the corresponding element.}
  \end{centering}
\end{figure}

Running the simulation with different refinement modes 
and meshes showed that the multi-mesh configuration generally results in
a smaller problem, faster calculation, and better or similar error convergence
compared to the single-mesh configuration.
This is well illustrated in Fig.~\ref{fig:singlemultidof} 
and Fig.~\ref{fig:singlemulticpu}.
The result can be understood from Fig.~\ref{fig:comsol-conc-volt} --- in the boundary
regions $\partial \Omega_1$ and $\partial\Omega_3$ the concentration gradient
is greater than voltage gradient, namely $\nabla C >> \nabla \phi$. Therefore
refining the mesh for both variables is not reasonable in terms of
number of degrees of freedom. For instance,
the solution space with corresponding mesh in case of
HP\_ANISO at $t=0.7\ s$ is shown in Fig.~\ref{fig:cphi}. The corresponding polynomial
degree space is shown in Fig.~\ref{fig:poly}. Notice that the adaptive algorithm
has increased the maximum polynomial degree for $C$ space to 7, however,
at the same time, the maximum polynomial degree for the $\phi$ space is one. Furthermore,
the mesh is significantly more refined for $C$.
For most cases using the multi-mesh results in similar or better CPU time, i.e.
multi-mesh takes less computational resources. Only excpetion was HP\_ANISO\_H 
refinement mode for which the single mesh configuration resulted slightly faster
calculation time. 
Based on the results, only multi-mesh configurations will be considered
in the following comparisons.

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoaniso_dof}
  \caption{\label{fig:isoanisodof} NDOFs in case 
  of multi-mesh solutions with H\_ISO, H\_ANISO,
  HP\_ISO, HP\_ANISO, and single-mesh solution with HP\_ANISO\_H
  refinement modes. (Notice log Y scale)}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoaniso_cpu}
  \caption{\label{fig:isoanisocpu} CPU times in case 
  of multi-mesh solutions with H\_ISO, H\_ANISO,
  HP\_ISO, HP\_ANISO, and single-mesh solution with HP\_ANISO\_H
  refinement modes. (notice log Y scale)}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoanisop_dof}
  \caption{\label{fig:isoanisopdof} NDOFs in case 
  of multi-mesh solutions with P\_ISO, P\_ANISO, and
  HP\_ANISO\_P refinement modes.}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoanisop_cpu}
  \caption{\label{fig:isoanisopcpu} CPU times in case 
  of multi-mesh solutions with P\_ISO, P\_ANISO, and
  HP\_ANISO\_P refinement modes.}
  \end{centering}
\end{figure}
To narrow down the list of refinement modes for the given problem, first the
isotropic and anisotropic adaptivities were compared. The \emph{p}-adaptivity
modes were compared separately as they used more refined mesh (Fig.~\ref{fig:mesh}~(c)).
Fig.~\ref{fig:isoanisodof} and Fig.~\ref{fig:isoanisocpu} show the comparison
of H\_ISO, H\_ANISO, HP\_ISO, HP\_ANISO, and HP\_ANISO\_H modes in terms of CPU time
and problem size.
Fig.~\ref{fig:isoanisopdof} and Fig.~\ref{fig:isoanisopcpu} show the similar
comparsion for the P\_ISO, P\_ANISO, and HP\_ANISO\_P modes.
It can be clearly seen that the anisotropic refinement modes result in a reasonable problem
size and the problems solve within a reasonable calculation time. It is interesting
to note that HP\_ANISO results in the smallest problem size. Also, in
\emph{p}-adaptivity group, HP\_ANISO\_P results in the smallest problem size
at each time step, whereas P\_ISO and P\_ANISO have a very largy problem size
during the first time steps of the solution.
Here the term ``reasonable problem size''
means that the number of degrees of freedom in time converges
to so that $N_{dof}<500$, and the term ``reasonable calculation time''
means that the calculation (step $\tau=0.01\ s$, physical
time $t_{end}=3.0\ s$) time $t$ on a givem system was $t<500\ s$.
Although these parameters are empirical, they serve as an upper limit, given
that the most refinement modes give significantly smaller results:
$t<<500\ s$ and $N_{dof} << 500$.

@@

Based on the results in the previous subsection, H\_ANISO, HP\_ANISO,
and HP\_ANISO\_H from the \emph{hp/h}-adaptivity group and HP\_ANISO\_P
from \emph{p}-adaptivity group will be compared
in terms of problem size and cumulative CPU time. 
In all of the cases, the relative 
error at each time step remained below
the threshold which were set to $e_{th}=0.5\%$ between the coarse mesh
and fine mesh solutions, therefore the error-time plot will not be considered.

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{cpu}
  \caption{\label{fig:cpu} Comparative CPU time for different refinement modes.}
  \end{centering}
\end{figure}

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{dof}
  \caption{\label{fig:dof} Comparative NDOFs at each time step for 
  different refinement modes.}
  \end{centering}
\end{figure}

Fig.~\ref{fig:cpu} shows the cumulative CPU time for different refinement 
modes at each time step. All the calculations were done on the same computer.
Here we see that HP\_ANISO\_H and HP\_ANISO require the most
resources which can be understood from the fact that these 
refinement modes have the largest number of 
candidates (see Section.~\ref{sec:model}) from which the refinement method is chosen.
At the same time, HP\_ANISO\_P is the fastest among all the other refinement modes.
Fig.~\ref{fig:dof} shows the NDOFs at each time step.
It can be seen that the HP\_ANISO results in the 
smallest problem size --- $N_{dof} \approx 125$. 
All the other refinement modes result in a 
problem size of approximately ($N_{dof} \approx 250$). So we have
two notable refinement modes --- HP\_ANISO\_P because of the fast
solution, and HP\_ANISO because of the small problem size. In the following
subsection, optimization of those two modes will be considered.

@@

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{refined_cpu}
  \caption{\label{fig:refined-cpu} Cumulative CPU time for HP\_ANISO and HP\_ANISO\_P
  with different initial meshes.}
  \end{centering}
\end{figure}

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{refined_dof}
  \caption{\label{fig:refined-dof} NDOFs at each time step for
  HP\_ANISO and HP\_ANISO\_P with different meshes.}
  \end{centering}
\end{figure}

So far we have seen that HP\_ANISO results in the smallest problem size, but requires
quite a lot CPU time compared to HP\_ANISO\_P that is the fastest.
When it comes to a large domain or 3D modeling the problem
size becomes the most important factor. Therefore, we consider HP\_ANISO the most
suitable refinement mode to the given problem. Thus some ways to optimize the time
factor will be considered. The desirable output would be HP\_ANISO problem size
close to HP\_ANISO\_P CPU time. One way to optimize the problem
is to choose somewhat more refined initial mesh. Other way one could think
to optimize the problem would be to change
the refinement frequency during the solving process. Recall that up to this point,
the initial mesh has been loaded in the beginning of each time step. True,
by employing the optimizations, we already must know something about the problem
and its solution beforehand. However, it could still be practial when solving a real problem
in a large domain.

@@

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{unreffreq_dof}
  \caption{\label{fig:unreffreq-dof} NDOFs at each time step for
  HP\_ANISO and HP\_ANISO\_P with mesh unrefinement at each time step and at over each
  	time step after first $0.5\ s$ physical solution time.}
  \end{centering}
\end{figure}
The next proposed optimization involved changing the unrefinement frequency.
It is known that the concentration gradient $\nabla C$ changes the most in the initial phase
of the calculation, therefore, the unrefinement after each time step was performed
until $t=0.5\ s$ (physical time). After that, the
unrefinement was performed in $\Delta t = 0.10\ s$ interval.
However, this optimization did not result in a stable solution, i.e. the problem size 
does not remain steady, but starts to oscillate depending on the unrefinement
frequency. This is shown in 
Fig.~\ref{fig:unreffreq-dof}.

@@

\subsection{More general results}
\begin{figure}
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{cphiorders}
  \caption{\label{fig:cphi-orders} Solutions $C$ and $\phi$
  and corresponding polynomial degrees of the elements at
  $t=0.7\ s$. HP\_ANISO refinement mode was used. The height
  in the solution graphs indicates the value.}
  \end{centering}
\end{figure}

@@

\emph{
PAVEL: Should I try to include the similar solution from comsol and do
some comparison here? I got this idea just before commiting the final
draft to you?}

>>>>

\section{Numerical Results and Comparisons}\label{sec:results}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=.8\columnwidth]{mesh}
  \caption{\label{fig:mesh} Initial coarse mesh (a),
  	half refined mesh (b) and refined mesh (c). The coarse mesh
	and refined mesh were used in the initial calculations, the latter one
	in case of \emph{p}-adaptivity (including HP\_ANISO\_P). The half-refined mesh was
	used later to optimize the \emph{hp}-adaptive refinement solutions.}
  \end{centering}
\end{figure}
Calculation for all the refinement modes were performed
in both single-mesh and multi-mesh configurations. 
The following numerical results were recorded for each 
refinement mode: converged relative error, cumulative CPU
time, and the problem size in terms of number
of degrees of freedoms (NDOFs) at each time step. 
Two types of initial meshes were used --- in case of only \emph{p}-adaptivity,
more refined mesh was used (Fig.~\ref{fig:mesh}~(c)) to ensure
the error convergence.
When the element size refinement
was also enabled (all \emph{h}/\emph{hp} refinement modes), very coarse initial mesh
was used (see Fig.~\ref{fig:mesh}~(a)) to let the adaptivity
algorithm find the most optimal mesh. So coarse initial mesh might not be
suitable for most the practical applications, as will be demonstrated in the
end of this section, however, it provides a good insight into the
adaptivity performance of Hermes.
In both cases, the initial mesh was loaded at each time step of the
calculation.

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{cphi}
  \caption{\label{fig:cphi} Concentration $C$
  and voltage $\phi$ at two different time steps
  (HP\_ANISO refinement mode was used). This is a 2D solution shown
  as in 3D, where the height indicates the values of $C$ and $\phi$.}
  \end{centering}
\end{figure}
Example of the solution at $t=0.7\ s$ and $t=3.0\ s$ 
calculated with HP\_ANISO refinement mode is shown
in Fig.~\ref{fig:cphi}. The time $t=0.7\ s$ was chosen because
by that time (with given constants in Table~\ref{Table:used-constants}),
some ionic migration has already taken place and
the concentration gradients near the boundaries $\partial\Omega_1$ and
$\partial\Omega_3$ have formed. The automatic mesh refinements
at different time steps are clearly visible in the figure --- especially
near $\partial\Omega_1$, where the $C_{\Omega_1}\approx 0$ and
the gradient $\nabla C$ is
moving in time (this was also demonstrated in Fig.~\ref{fig:comsol-conc-volt}).

@@

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{singlemulti_dof}
  \caption{\label{fig:singlemultidof} NDOFs in case 
  of single-mesh and multi-mesh solutions with HP\_ISO
  and HP\_ANISO refinement modes. (Notice log Y scale)}
  \end{centering}
\end{figure}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{singlemulti_cpu}
  \caption{\label{fig:singlemulticpu} CPU times in case
  of single-mesh and multi-mesh solutions with HP\_ISO
  and HP\_ANISO refinement modes. (Notice log Y scale)}
  \end{centering}
\end{figure}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{poly}
  \caption{\label{fig:poly} Polynomial degree space
  for $C$ and $\phi$ at $t=0.7\ s$ The color indicates
  the maximum polynomial degree of the corresponding element.}
  \end{centering}
\end{figure}

Running the simulation with different refinement modes 
and meshes showed that the multi-mesh configuration generally results in
a smaller problem, faster calculation, and better or similar error convergence
compared to the single-mesh configuration.
This is well illustrated in Fig.~\ref{fig:singlemultidof} 
and Fig.~\ref{fig:singlemulticpu}.
The result can be understood when considering Fig.~\ref{fig:cphi} --- in the vicinity
of the boundaries $\partial \Omega_1$ and $\partial\Omega_3$ the concentration gradient
is greate than voltage gradient, namely $\nabla C >> \nabla \phi$. Therefore,
adapting the mesh for both variables is not reasonable in terms of
number of degrees of freedom. For instance,
the solution space with corresponding mesh in case of
HP\_ANISO at $t=0.7\ s$ is shown in Fig.~\ref{fig:cphi}. The corresponding polynomial
degree space for multi-mesh configuration is shown in Fig.~\ref{fig:poly}. 
Notice that the adaptivity algorithm
of Hermes has increased the maximum polynomial degree for the $C$-space to~7, however,
the maximum polynomial degree for the $\phi$-space is~2. Furthermore,
the mesh is significantly more refined for $C$.
For most of the cases using the multi-mesh results in similar or better CPU time.
It means that multi-mesh takes less computational resources. Only excpetion was HP\_ANISO\_H 
refinement mode for which the single mesh configuration resulted slightly faster
calculation time. 
Based on the results, only multi-mesh configurations will be considered
in the following comparisons.

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoaniso_dof}
  \caption{\label{fig:isoanisodof} NDOFs in case 
  of multi-mesh configuration with H\_ISO, H\_ANISO,
  HP\_ISO, HP\_ANISO, and single-mesh configuration with HP\_ANISO\_H
  refinement modes. (Notice log Y scale)}
  \end{centering}
\end{figure}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoaniso_cpu}
  \caption{\label{fig:isoanisocpu} CPU times in case 
  of multi-mesh configuration with H\_ISO, H\_ANISO,
  HP\_ISO, HP\_ANISO, and single-mesh configuration with HP\_ANISO\_H
  refinement modes. (notice log Y scale)}
  \end{centering}
\end{figure}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoanisop_dof}
  \caption{\label{fig:isoanisopdof} NDOFs in case 
  of multi-mesh configuration with P\_ISO, P\_ANISO, and
  HP\_ANISO\_P refinement modes.}
  \end{centering}
\end{figure}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoanisop_cpu}
  \caption{\label{fig:isoanisopcpu} CPU times in case 
  of multi-mesh configuration with P\_ISO, P\_ANISO, and
  HP\_ANISO\_P refinement modes.}
  \end{centering}
\end{figure}
To narrow down the list of refinement modes for the given problem, first the
isotropic and anisotropic adaptivities were compared. The \emph{p}-adaptivity
modes were compared separately as they used more refined initial mesh 
(shown in Fig.~\ref{fig:mesh}~(c)).
Fig.~\ref{fig:isoanisodof} and Fig.~\ref{fig:isoanisocpu} provide the comparison
of H\_ISO, H\_ANISO, HP\_ISO, HP\_ANISO, and HP\_ANISO\_H modes in terms of CPU time
and NDOFs.
Fig.~\ref{fig:isoanisopdof} and Fig.~\ref{fig:isoanisopcpu} show the similar
comparsion for the P\_ISO, P\_ANISO, and HP\_ANISO\_P modes.
It is not difficult to see that the anisotropic refinement modes result in a reasonable problem
size and the problems solve within a reasonable calculation time. It is interesting
to note that HP\_ANISO results in the smallest problem size. Also, in
\emph{p}-adaptivity group, HP\_ANISO\_P results in the smallest problem size
at each time step, whereas P\_ISO and P\_ANISO result in a very large problem size
during the first time steps of the solution.
Here the term ``reasonable problem size''
means that the number of degrees of freedom in time converges
to so that $N_{DOF}<500$, and the term ``reasonable calculation time''
means that the calculation (step $\tau=0.01\ s$, physical
time $t_{end}=3.0\ s$) time $t$ on a givem system was $t<500\ s$.
Although these parameters are empirical, they serve as an upper limit, given
that the most refinement modes give significantly better results, i.e.
$t<<500\ s$ and $N_{DOF} << 500$.

@@

Based on the results in the previous subsection, H\_ANISO, HP\_ANISO,
and HP\_ANISO\_H from the \emph{hp/h}-adaptivity group and HP\_ANISO\_P
from \emph{p}-adaptivity group will be compared
in terms of NDOFs and cumulative CPU time. 
In all of the cases, the relative 
error at each time step remained below
the threshold which was set to $e_{th}=0.5\%$ between the coarse mesh
and fine mesh solutions, therefore the error-time plot will not be considered.

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{cpu}
  \caption{\label{fig:cpu} Comparative CPU time for different refinement modes.}
  \end{centering}
\end{figure}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{dof}
  \caption{\label{fig:dof} Comparative NDOFs at each time step for 
  different refinement modes.}
  \end{centering}
\end{figure}

Fig.~\ref{fig:cpu} shows the cumulative CPU time for different refinement 
modes at each time step --- the results were recorded on the same computer
for the CPU times to be comparable.
We see that HP\_ANISO\_H and HP\_ANISO require the most
resources. This can be understood from the fact that in case of these 
refinement modes the adaptivity algorithm has the largest number of 
candidates (see Section.~\ref{sec:model} and Fig.~\ref{fig:refinements})
from which a particular refinement can be chosen.
At the same time, HP\_ANISO\_P is the fastest compared to the other refinement modes.
This can be attributed to the finer inital mesh.

Fig.~\ref{fig:dof} shows the NDOFs at each time step.
It can be seen that the HP\_ANISO results in the 
smallest problem size --- $N_{DOF} \approx 125$. 
All the other refinement modes result in a 
problem size of approximately ($N_{DOF} \approx 250$). Therefore
the HP\_ANISO is the most efficient in terms of memory consumption.

As we have filtered out
two notable refinement modes for the given problem --- HP\_ANISO\_P because of the fast
calculation, and HP\_ANISO because of the small problem size. In the following
subsection, some optimizations to reduce the CPU time of
HP\_ANISO and reduce the NDOFs of HP\_ANISO\_P will be considered.

@@

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{refined_cpu}
  \caption{\label{fig:refined-cpu} Cumulative CPU time for HP\_ANISO and HP\_ANISO\_P
  with different initial meshes.}
  \end{centering}
\end{figure}

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{refined_dof}
  \caption{\label{fig:refined-dof} NDOFs at each time step for
  HP\_ANISO and HP\_ANISO\_P with different meshes.}
  \end{centering}
\end{figure}

When it comes to a large domain or 3D modeling, the problem
size becomes a very important factor. Namely, 3D full scale solutions tend to use
a lot of memory. Therefore, we consider HP\_ANISO the most
suitable refinement mode for the given problem. 
Hence, some ways to optimize the model based on HP\_ANISO
mode to improve the CPU time
%TODO, continue here
factor without significantly compromising the NDOFs will be considered. 
The desirable output would be HP\_ANISO problem size
close to HP\_ANISO\_P CPU time. One way to optimize the problem
is to choose more refined initial mesh. Another way would be to change
the refinement frequency during the solving process. In theory, this
could result in a fewer adaptivity steps which are quite expensive in terms
of CPU time. Recall that up to this point,
the initial mesh has been loaded in the beginning of each time step. 
True, by employing these optimizations, we must know something about the problem
and its solution beforehand. However, it could still be practial when solving a real problem
in a large domain.

@@

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{unreffreq_dof}
  \caption{\label{fig:unreffreq-dof} NDOFs at each time step for
  HP\_ANISO and HP\_ANISO\_P with mesh unrefinement at each time step and at over each
  	time step after first $0.5\ s$ physical solution time.}
  \end{centering}
\end{figure}
The next proposed optimization involved changing the unrefinement frequency.
It is known that the concentration gradient $\nabla C$ changes the most in the initial phase
of the calculation, therefore, the unrefinement after each time step was performed
until $t=0.5\ s$ (physical time). After that, the
unrefinement was performed in $\Delta t = 0.10\ s$ interval.
However, this optimization did not result in a stable solution, i.e. the problem size 
does not remain steady, but starts to oscillate depending on the unrefinement
frequency. This is shown in 
Fig.~\ref{fig:unreffreq-dof}.

@@

\subsection{More general results}
\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{cphiorders}
  \caption{\label{fig:cphi-orders} Solutions $C$ and $\phi$
  and corresponding polynomial degrees of the elements at
  $t=0.7\ s$. HP\_ANISO refinement mode was used. The height
  in the solution graphs indicates the value.}
  \end{centering}
\end{figure}

@@
